# Results

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

# clear environment
rm(list=ls())

# load packages
library(bookdown)
library(tinytex)
library(tidyverse)
library(googlesheets4)
library(lubridate)
library(readr)
library(readxl)
library(writexl)
library(hms)
library(DT)
library(xlsx)
library(leaflet)
library(DT)
library(ggpubr)
library(plotrix)
library(packrat)
library(foreign)
library(kableExtra)

# set plotting themes

## geom_col plots theme
col_theme <- theme(axis.title = element_text(size = 14, face = "bold"),
                   strip.text = element_text(size = 14, face = "bold"),
                   legend.title = element_text(size = 14, face = "bold"),
                   legend.text = element_text(size = 14),
                   axis.text = element_text(size = 14))

## geom_points plots theme
points_theme <- theme(axis.title = element_text(size = 14, face = "bold"),
                   strip.text = element_text(size = 14, face = "bold"),
                   legend.title = element_text(size = 14, face = "bold"),
                   legend.text = element_text(size = 14),
                   axis.text = element_text(size = 11, face = "bold"),
                   title = element_text(size = 18))

# function to exclude multiple items per column
'%ni%' <- Negate('%in%')

# clarify select function
select <- dplyr::select
```

```{r, echo = F}


# notes 

## total number of sites with anad species ("n of x sites had newly id'd spp")

## 

```

```{r, echo = F}
# summarise total capture effort

## read in 
url <- "https://docs.google.com/spreadsheets/d/1S0hwY4EQo9Xtz5d4UnSBCbZoJ0ft_HGQrarfmA5dwNY/edit#gid=0"

## prep data sources

### sample events
sample_events <- read_sheet(url, sheet =  "A_Sample_Event") %>%
  # retain only TU events
  filter(project_org == "TU") %>%
  transform(site_arrive_time = as_hms(site_arrive_time),
            site_depart_time = as_hms(site_depart_time)) %>%
  select(-QC1,-QC2,-data_entry,-notes,-crew)

### fish count
fish_dat <- read_sheet(url, sheet = "F_Fish_ID") %>%
  select(-QC1,-QC2,-data_entry)  %>%
  
  # retain only TU events by using an inner join to only dates when TU events occurred. Assumes that no other simultaneous fieldwork was done elsewhere that day
  inner_join(sample_events, by = c("site_depart_date","site"))



### how many unique sampling events (unique date/site) did we have in 2021?
n_events <- sample_events%>%
  group_by(site,site_arrive_time) %>%
  tally() %>%
  nrow() %>%
  as.character()

### how many unique event days did we have in 2021?
n_days <- sample_events%>%
  select(site_arrive_date) %>%
  distinct() %>%
  nrow() %>%
  as.character()


### number of sites

#### total value only
n_sites_total <- sample_events %>%
  group_by(site) %>%
  tally() %>%
  nrow() %>%
  as.character()

#### table of values by site
n_sites <- sample_events %>%
  group_by(site) %>%
  tally() 

### total fish
fish_ct <- fish_dat %>%
  summarise(n = sum(count)) %>%
  as.character()

### total unique species
spp_ct <- fish_dat %>%
  group_by(species) %>%
  summarise(n = sum(count)) %>%
  nrow() %>%
  as.character()

### count by species
spp_ct_n <- fish_dat %>%
  group_by(species) %>%
  summarise(n = sum(count))

### count by species at each site
site_spp_ct <- fish_dat %>%
  group_by(species,site) %>%
  summarise(n = sum(count)) %>%
  pivot_wider(names_from = "species", values_from = "n") 

### total fish at each site
total_by_site <- fish_dat %>%
  group_by(site) %>%
  summarise(total_fish = sum(count))

#### join site totals to species totals by site
z <- left_join(site_spp_ct,total_by_site)

#### remove NAs
site_spp_ct[is.na(site_spp_ct)] <- 0

```

*Note: All fishing effort and capture results are current as of `r Sys.Date()`.*

## Fish capture

Since initiation of fieldwork in summer 2021:

-   Over course of `r n_days` fieldwork days in summer 2021, KPTU in collaboration with KWF have conducted a total of `r n_events` sampling events at `r n_sites_total` unique sites, including the training event. See the interactive project map in the Methods section ("[Site selection](##%20Site%20selection)") for a current map of where and when sampling events have occurred.

-   Table \@ref(tab:site-visits) summarises total samping events by individual site. We typically sampled a site only one time each in summer 2021.

-   We have captured `r fish_ct` fish comprised of `r spp_ct` unique species. Table \@ref(tab:ct) summarizes current fish capture count by species.

-   Table \@ref(tab:site-spp-ct) summarizes total fish capture for each species by site.

```{r, site-visits, echo = F, fig.cap = "cap"}
n_sites %>%
  kbl(caption = "Total site visits") %>%
  kable_paper(full_width = F, html_font = "Cambria")

#, 
 #             bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# fix w/ webshot or make non-html
```

***

<br>

```{r ct, echo = F, fig.cap = "Total fish capture count by species"}

spp_ct_n %>%
  kbl(caption = "Total fish capture count by species") %>%
  kable_paper(full_width = F, html_font = "Cambria")
#, 
 #             bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

***

<br>

```{r site-spp-ct, echo = F, fig.cap = "cap"}

site_spp_ct %>%
  kbl(caption = "Total fish capture count by species at each site.") %>%
  kable_paper(full_width = F, html_font = "Cambria") 

#%>%
#, 
 #             bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  #kable_styling(latex_options = c("scale_down","striped")) %>%
#  scroll_box(width = "100%", height = "500px")

```

<br>

## Aquatic Resource Permit and AWC Nominations

2021 fish capture data was reformatted in to a csv file for submission to ADF&G in fulfillment of the required [Aquatic Resource Permit](http://www.adfg.alaska.gov/index.cfm?adfg=otherlicense.aquatic_reports)^[http://www.adfg.alaska.gov/index.cfm?adfg=otherlicense.aquatic_reports]; available in the online in [this project's GitHub repo](https://github.com/Kenai-Watershed-Forum/tu_awc_expansion/tree/main/output)^[https://github.com/Kenai-Watershed-Forum/tu_awc_expansion/tree/main/output]. These data will be also be shared with the local Area Biologist at the Kenai River Center to determine the best strategy for submitting nomination data in January 2022.

The total number of stream miles / lake acres of successfully nominated to the AWC will be posted here when available in January 2022.

```{r, echo = F}

# prep data format for ARP report: https://www.adfg.alaska.gov/sf/SARR/AWC/index.cfm?ADFG=nomSubmit.about

# get unique sampling events (site, location, fishing method)
sample_effort <- read_sheet(url, sheet = "C_Sample_Effort") %>%
  filter(!is.na(gear_type)) %>%
  select(site,site_depart_date, gear_type) %>%
  distinct()

# join sample effort data
fish_dat <- left_join(fish_dat,sample_effort, by = c("site","site_depart_date"))

# prepare data into ARP format
arp <- fish_dat %>%
  
  # rectify ARP column named "additional count"; leave blank if count = 1, make n-1 if count > 1
  # presumed that count = 1 if additional count = 0
  mutate(`Additional count (1)` = ifelse(fish_dat$count <= 1, "0",fish_dat$count - 1)) %>%
  
  # remove extraneous columns
  select(-site_arrive_time, -site_depart_time,-photo_id, -count, -camera_id, -project_org,
         -site_arrive_date) %>%
  
  # rename existing columns
  rename(`Latitude (decimal degrees)` = latitude,
         `Longitude (decimal degrees)` = longitude,
         Date = site_depart_date,
         `Location ID (optional)` = site,
         Species = species,
         `Life stage` = lifestage,
         `Weight (g)` = wt_g,
         `Length (mm) (NO estimates or ranges)` = fl_mm,
         `Length method` = length_method,
         `Disposition (1)` = disposition,
         Comments = notes,
         `Fish collection method` = gear_type) %>%
  
  # remove extraneous columns
  select(-observer_name) %>%
  
  # create missing columns
  mutate(Datum = "",
         `Coordinate determination method` = "",
         `Name of water body` = fish_dat$site,
         `Observer name (the first and last name of the person handling fish)` = fish_dat$observer_name,
         `Age method (sample type)` = "",
         `ADF&G GCL` = "",
         `Additional count (2)` = "",
  `Disposition (2)` = "",
  `Anesthesia/ Sedative Used (Leave blank if none used)` = "",
  Sex = "",
  Age = ""
         ) %>%
  
  select(
  `Location ID (optional)`,
  `Latitude (decimal degrees)`,
  `Longitude (decimal degrees)`,
  Datum,
  `Coordinate determination method`,
  `Name of water body`,
  Date,
  `Observer name (the first and last name of the person handling fish)`,
  `Fish collection method`,
  `Species`,
  `Life stage`,
  `Length (mm) (NO estimates or ranges)`,
  `Length method`,
  `Weight (g)`,
  Sex,
  Age,
  `Age method (sample type)`,
  `ADF&G GCL`,
  `Additional count (1)`,
  `Disposition (1)`,
  `Additional count (2)`,
  `Disposition (2)`,
  `Anesthesia/ Sedative Used (Leave blank if none used)`,
  Comments						
  )

# export csv to repo
write.csv(arp, "output/aquatic_resource_permit_2021.csv", row.names = F)


```



